% !TEX root = ../../thesis.tex

\section{Introduction}\label{sec:introduction}

Cyber-physical systems (CPS) that consist of several actuators, sensors, controllers, and communication networks, are becoming increasingly prevalent in many infrastructures. Securing these systems against malicious attacks and communication failures is an important problem \cite{cps1}. Recently, several aspects of the problem of securing complex CPS have been investigated, e.g., the networking security among cyber devices \cite{security_0}-\!\!\cite{comm_net_4} and the early detection of attacks \cite{comm_early_1}-\!\!\cite{comm_early_2}. Researchers have also tried to understand how we can securely estimate the state of a dynamical system from a set of corrupted sensor measurements by leveraging the system dynamics \cite{ctrl_sec8}.



Several researchers have focused on linear dynamical systems \cite{ctrl_sec8}-\!\!\cite{new4}, and have tried to provide security guarantees. The existing literature related to secure state estimation in linear dynamical systems can be broadly categorized into two classes depending on the noise model for sensor measurements: 1) noiseless measurements, and 2) noisy measurements. For the noiseless framework, the studies in \cite{cps1}, \cite{ctrl_sec8}, \cite{new2} show that, sensor attacks can be detected and corrected under certain conditions, and hence the state of the system can be accurately reconstructed. Fawzi \textit{et al.} \cite{ctrl_sec8} focus on secure estimation and control of linear time-invariant systems, and assume that the set of attacked nodes does not change over time. The authors then formulate the system under attack as an estimation problem without any limiting assumption on attack signals, and propose a novel method for error estimation and correction. The main drawback of the study is that the set of attacked nodes is assumed to be fixed.


\textcolor{black}{Hu \textit{et al.} \cite{ctrl_sec9} extends} the results in \cite{ctrl_sec8} to scenarios in which the set of attacked nodes can change over time, and show that under a certain condition, the secure estimation problem with time-varying attacked nodes is equivalent to the classical error correction problem. The authors provide a novel method to guarantee accurate decoding, and then propose a secure estimation method which is a combination of the proposed secure estimator and the Kalman Filter (KF). Finally, they demonstrate the performance of their algorithm through numerous simulations. The problem of distinguishing between measurement noise and attack signals arises when the sensor measurements are affected by both noise and attack signals. This problem is studied in \cite{new4}-\!\!\cite{new10}. The authors provide sufficient conditions under which the
sparse attack vectors can be distinguished from measurement noise.


\textcolor{black}{Although many approaches in the literature are addressing the secure state estimation problem, they are based
on linear dynamical systems. Hence, the existing secure estimators can be applied to nonlinear dynamical systems if we linearize these systems. It is well known that the linearization of nonlinear dynamical systems can result in the following drawbacks:
\begin{enumerate}
\item Linearization is reliable if the higher order terms in the Taylor series expansion can be eliminated; otherwise, the linearized model may perform poorly. %Therefore, the linearized model is only valid under small perturbation.
\item \textcolor{black}{Linearization can be applied when all the eigenvalues of the Jacobian matrix have nonzero real part. However, this is not always the case.}
\end{enumerate}
For example, linearized power system models are only valid under small perturbations in the system at hand. Under a severe disturbance, such as a single or multi-phase short-circuit or a generator loss, the linearized model does not remain valid \cite{Kundur}-\!\!\cite{nonlin_est}. Therefore, the existing techniques lack performance guarantees when the system undergoes large perturbations which are typical of highly loaded practical systems. To overcome the above drawbacks, we develop a secure state estimation method without linearization or calculation of Jacobian matrices.} \textcolor{black}{Note that feedback linearization techniques transform the nonlinear system into an equivalent linear system through a change of variables and a suitable control input. Even with such techniques, the secure estimation problem for nonlinear dynamical systems is a nonlinear problem.}



\textcolor{black}{In this study, we investigate the secure estimation of the state of a nonlinear dynamical system from a set of corrupted measurements without any assumption on the sensor attacks or corruptions (i.e., corruptions can follow any particular model). Our only assumption concerning the corrupted sensors is about the number of sensors that are corrupted due to attacks or failures. We consider two classes of nonlinear systems, and design secure state estimators for \textcolor{black}{these} assuming that the set of attacked sensors can change with time. \textcolor{black}{A practical example of such a cyber attack is described in \cite{kundur_time}, where a multi-switch attack, in which different switches in a power network are attacked at different times, is designed to lead to stealthy and wide-scale cascading failures in the power system.}
%\textcolor{black}{This is possible in practice. For example,} in power systems, the set of attacked sensors can change with time \cite{kundur_time}.}
We then propose a technique which enables us to transform the nonlinear dynamics into a set of linear equations, and apply the classical error correction method to the equivalent linear system. The proposed secure state estimators are computationally efficient and can be solved exactly without iteration. \textcolor{black}{In addition, our estimator relies on the observability of the transformed linear system, which is much simpler to check than verifying the observability of nonlinear systems.}} %\textcolor{black}{(This seems a bit abrupt as we do not mention checking observability of nonlinear systems before this point?)}






\textcolor{black}{The work closest to ours is \cite{shoukry} in which Shoukry \textit{et al.} studied differentially flat nonlinear systems under sensor attack and assumed that the set of attacked sensors do not change with time. Using $s$-sparse observability for nonlinear systems, the authors proposed a combinatorial estimator, and an iterative satisfiability modulo theory-based algorithm to solve the resulting combinatorial estimation problem. However, it may be hard to check the observability of nonlinear systems, and the assumption of fixed attacked nodes may be restrictive.}



%\textcolor{black}{\st{We provide guarantees on the achievable state estimation error against arbitrary corruptions, and characterize the number of errors that can be perfectly corrected by an estimator. (I think we removed these)}} \textcolor{black}{Agree}


To illustrate how the proposed nonlinear secure estimation approach can be applied to practical systems, we focus on an interconnected power system comprising several synchronous generators, transmission lines, buses, and energy storage units. We assume that all the physical devices are controlled via a wide area control system (WACS) as well as local controllers, and that these control systems use the synchrophasor technology, phasor measurement units (PMU), to maintain the system's stability\footnote{The secondary generation control in power systems is an example of such cyber-physical structures. In this system, measurements and control signals are telemetered to and from the generating units and that control center adjusts the set-point of each generator based upon the integral of the frequency error.}. The WACS and local controllers employ advanced data acquisition, communications, and control to enable increased efficiency and reliability of power delivery \cite{pmu_w_0}-\!\!\cite{wacs_ref8}. Several methods for power system state estimation have been proposed \cite{ref_v11}-\!\!\cite{ref_v16}. %Linearization and Jacobian matrix calculation are indispensable in these methods.
\textcolor{black}{All these methods rely on the linearization.} To overcome the drawbacks of linearization, Wang \textit{et al.} \cite{nonlin_est} develop a dynamic state estimation method that requires neither linearization nor calculation of Jacobian matrices. However, the authors only consider Gaussian noise. Extensive work has been done on monitoring and autonomous feedback control for WACS \cite{wacs_ref8}, and on secure state estimation of static states \cite{ref_v1}-\!\!\cite{ref_v2}. However, these works have not studied how to identify cyber-physical attacks or communication failures when dynamic states such as generator' phase angles are estimated, and how to perform secure state estimation (dynamic state estimation) for the WACS.

We focus on secure estimation for the wide area control system of the power network assuming that the installed PMUs at different generator buses are connected through a communication network which sends PMU measurements to the WACS as well as the local controllers in the power network. We assume that the communication channels from the WACS to the generators are secured while other channels and PMUs are not secured and are subject to cyber attacks and failures. Therefore, the WACS needs to perform secure state estimation to reconstruct the system's states before using the received data for computing wide area control signals, and to monitor the operation of local controllers. By using the developed secure estimation technique, we propose a secure state estimator for the wide area control of the power system, and numerically show that the proposed algorithm significantly improves the performance of the cyber layer in power systems.

The paper is organized as follows: In Section \ref{sec:review}, we review the classical error correction methods. In Section \ref{sec:nonlinear_estimation}, we formulate the nonlinear state estimation problem and propose a solution technique for two classes of nonlinear systems. We then illustrate how the proposed secure state estimation approach can be applied to power systems in Section \ref{sec:application} and \ref{sec:application1}. Finally, in Section \ref{sec:example}, we numerically demonstrate the effectiveness of the proposed secure estimation algorithm.