\documentclass[../../thesis.tex]{subfiles}
 

 
\begin{document}

%\input{chapters/building_model/main}
%\input{chapters/building_exp/main}


%Increased penetration of renewable energy sources has increased the demand for frequency regulation reserves.
%
%%More specifically, by adjusting the electricity consumption of the supply fans of the HVAC system. 
%A main challenge in this application is that commercial buildings and their HVAC systems are often complex and subject to large disturbances such as occupancy.
%%Here, we describe a data-driven method to identify a model that describes the building's thermal dynamics and captures such disturbances.
%
%In this chapter, we demonstrate experimentally that commercial buildings equipped with VAV HVAC systems can provide frequency regulation. 
%sIn addition, we propose a control scheme that adjusts the electricity consumption of the supply fans of the HVAC system to track the frequency regulation signal, and is suitable for scenarios where an accurate fan model is unavailable. %, e.g. buildings with large uncertainties. 
%We demonstrate the performance of our proposed control scheme through numerous field experiments conducted in an occupied commercial building, and in accordance with PJM's regulation market rules.


The operation of a reliable power grid depends on a balance of supply and demand at all times. 
%Any mismatch can cause the grid frequency to deviate from its nominal value (60 Hz in the U.S.).
Reserves known as ancillary services (AS) are used to correct any mismatch.
Amongst them, frequency regulation is the highest quality AS over which the grid operator has almost real-time control and is active continuously during normal operation of the grid, to maintain the grid frequency at its nominal value (60 Hz in the U.S.).
The recent rapid increase in the penetration of renewable energy sources has increased the volatility and uncertainty of electricity supply. This leads to a greater demand for frequency regulation reserves.
These reserves have been traditionally provided by conventional power generators. More recently, loads on the demand side have also been considered for this application.
% by increasing (and decreasing) their electricity consumption when the grid frequency increases (decreases).
In particular, commercial buildings are a tremendous untapped resource for frequency regulation. 
The feasibility of this proposal has bee investigated by numerous researchers through simulations and theoretical studies.
On the experimental side, there has only been a few field tests, and most of them were conducted in controlled laboratory environments with minimal uncertainties.

In this chapter, we demonstrate experimentally that commercial buildings equip-ped with HVAC systems can provide frequency regulation. First, we propose a control scheme that adjusts the electricity consumption of the supply fans of the HVAC system to track the frequency regulation signal.% and is suitable for scenarios where an accurate fan model is unavailable. %, e.g. buildings with large uncertainties. 
Then, we demonstrate the performance of our proposed control scheme through numerous field experiments conducted in an occupied commercial building, and in accordance with PJM's regulation market rules.

\input{chapters/building_exp/introduction.tex}
\input{chapters/building_exp/problem_statement.tex}
\input{chapters/building_exp/model_id.tex}
\input{chapters/building_exp/control.tex}
\input{chapters/building_exp/results.tex}
\input{chapters/building_exp/conclusion.tex}

\end{document}