\documentclass[../../thesis.tex]{subfiles}
 

 
\begin{document}

%\input{chapters/building_model/main}

%\begin{abstract}
%Commercial buildings are responsible for a large fraction of energy consumption in developed countries, and therefore are targets of energy efficiency programs. Motivated by the large inherent thermal inertia of buildings, the power consumption can be flexibly scheduled without compromising occupant comfort. This temporal flexibility offers opportunities for energy savings and the provision of frequency regulation to support grid stability. To realize these goals, it is of prime importance to identify a realistic model for the temperature dynamics of a building. In this paper, we identify a low-dimensional data-driven model and a high-dimensional physics-based model for the same system at different spatial granularities and temporal seasons using experimental data collected from an entire floor of an office building on the University of California, Berkeley campus. We perform a quantitative comparison in terms of estimates of the inherent thermal gains due to occupancy, open-loop prediction accuracies, and closed-loop control schemes. %Under an energy efficient control scheme, we show that both models result in a similar optimal control strategy, and therefore 
%We conclude that data-driven models could serve as a substitution for highly complex physics-based models with an insignificant loss of prediction accuracy for many applications.
%\end{abstract}

%Commercial buildings are responsible for a large fraction of energy consumption in developed countries, and therefore are targets of energy efficiency programs. 
%%The inter-temporal consumption flexibility of buildings can be harnessed to improve the energy efficiency of buildings, or to provide ancillary service to the power grid.
%The large inherent thermal inertia of buildings means that its power consumption can be flexibly scheduled without compromising occupant comfort. 
%This temporal flexibility offers opportunities for energy savings and the provision of frequency regulation to support grid stability. 
% 
%
%To realize these goals, it is of prime importance to identify a realistic model for the temperature dynamics of a building.

According to \cite{Perez-Lombard:2008aa}, commercial buildings account for more than 35\% of the total electricity consumption in the U.S., with an upward trend. HVAC systems are a major source of this consumption \cite{USenergy:2017}. %, ~35% of energy use in commercial buildings is for heating, cooling and ventilation.
Nevertheless, their power consumption can be flexibly scheduled without compromising occupant comfort, due to the thermal capacity of buildings. 
As a result, commercial buildings' HVAC systems have become the focal point of research, with the goal of utilizing this source of consumption flexibility. From the point of view of energy efficiency, researchers have studied optimization of building control in order to minimize power consumption \cite{Siroky:2011aa, Parisio:2014aa}.  
More recently, it has been proposed to engage buildings in supporting the supply quality of electricity and the grid stability, by participating in the regulation of electricity's frequency \cite{Balandat:2014contractdesign, Lin:2015exp, Vrettos:2014aggregation, Baccino:2014aa}.


All of the above research activities are based on a valid mathematical model describing the thermal behavior of buildings.  
Such a model should be identified using actual experimental data and capture realistic disturbances. 
In addition, it should describe the building's temperature evolution with suitable accuracy and spatial granularity, without being too computationally expensive.
However, there are many challenges in identifying such a building model. 
First, buildings often have different types of spaces that are subject to very different uncertainties. 
Second, disturbances such as occupancy are difficult to predict \textit{a priori}. 
Third, buildings are often not sufficiently excited, as they must satisfy strict regulatory requirements during regular operation, which limits the type and duration of excitation experiments that can be conducted.  

In this chapter, we present two methods for the identification of a building model: a physics-based method and a data-driven method. 
We use both techniques to identify building models for the same testbed using experimental data collected from the building, and quantitatively compare the models using various metrics, including open-loop prediction accuracy and closed loop control strategies. 
\\
\\
\noindent
\textit{This chapter is an adaptation of the paper in \cite{Zhou:2017modelcomp}.}

\input{chapters/building_model/introduction.tex}
\input{chapters/building_model/preliminaries.tex}
\input{chapters/building_model/data_driven_model.tex}
\input{chapters/building_model/physics_based_model.tex}
\input{chapters/building_model/comparison.tex}
\input{chapters/building_model/conclusions.tex}

\end{document}